\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{Brief Article}
\author{The Author}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

The Global Adaptation IndexªA decision tool for prioritizing investments in adaptation Table of ContentsForward	3Our Leadership	3Council of Advisers	5Council of Scientific Advisers	5The Dimensions of the Global Adaptation Problems	6Executive Summary	7The Global Adaptation Institute	7GaInª in Practice	8The Global Adaptation Problem	9The Global Adaptation IndexTM	10Vulnerability and Readiness: The Readiness Matrix	11The Scientific Approach: The Global Adaptation Indexª Ð GaInª	13Full Report	14Introduction	14Audience & Use	15Limitations	16Results & Rankings	17Adaptation Opportunities	18GaInª Structure	21Vulnerability and Readiness: The Readiness Matrix	21Vulnerable Sectors	23Vulnerability Components & Indicators 	25Vulnerability Indicators Overview	26Biophysical Exposure	27Socioeconomic Sensitivity	29Socioeconomic Adaptive Capacity	29Factors of Readiness	30Readiness Components & Indicators	31Country Profiles	32Future Work	33Adaptometerª	33Scaling to the local	34Linking to demonstration projects	34Version 2.0	34Methodology 	35Glossary of Terms	38References	40?ForwardOur Leadership Message from the Chairman, Council of Advisers It is my honor to support and guide this timely and important endeavor.  I decided to join this effort because I believe in the ability of humanity to innovate, create and implement practical solutions to the worldÕs most complex problems. I have also joined because as former President of Spain, I have learned that governments need to be able to remove obstacles for people to take destiny into their own hands. Our message is positive. Not apocalyptic. We believe that by creating the right incentives while expanding our knowledge and understanding of the need for adaptation, we can save lives and improve the livelihoods of people around the world, particularly those in areas most vulnerable to the effects of climate change and other global forces. I believe that the Global Adaptation Indexª will become a natural navigation chart for both the private and public sectors to direct adaptation investments efficiently and effectively. Because, in the end, our efforts are not truly compassionate unless they really work.SignatureThe Honorable Jose Maria AznarChairman, Council of Advisers, Global Adaptation Institute Former President of Spain (1996 Ð 2004)Message from the Chairman, Board of Directors As the global community struggles to reach agreement on an international climate mitigation policy, we donÕt have the luxury of postponing crucial investments in our food, water and energy systems as well as building protections for the hundreds of millions living in vulnerable coastal communities. While there has been an increase in discussions on adaptation in many international arenas, there has been little concrete action. The Global Adaptation Indexª is a groundbreaking tool that will guide capital to the sectors and countries where meaningful impact can be made in the area of adaptation. Too much time and money have been wasted on efforts that have small impacts or are undone by local institutional weaknesses. Starting now, we estimate that US$150-200 billion annually will be needed for the world to adapt to global changes. It is critical that those funds be directed where they will have the largest and most immediate impact.The Index, along with the Institute's important work in funding adaptation projects will help raise awareness of the need for global adaptation as well as mobilize private sector engagement in creating adaptation solutions. Mr. Kenneth A. Hersh Chairman, Board of Directors, Global Adaptation InstituteChief Executive Officer of NGP Energy Capital ManagementMessage from the Founding CEO [Text]SignatureDr. Juan Jose DaboubFounding CEO, Global Adaptation Institute Former Managing Director, World Bank Council of Advisers			 The Honorable Jose Maria AznarChairman of the Advisory Council, Global Adaptation InstituteMr. Aznar is the former President of Spain (1996 -2004)Anthony MorrisFounder, The Morris CompanyAna PalacioForeign Minister of Spain (2002-2004)Jorge QuirogaPresident of Bolivia (2001-2002)Vice President of Bolivia (1997-2001)Andreas WidmerCo-Founder, Seven Fund Council of Scientific Advisers			 Dr. Maria de Lourdes Dieck-AssadDirector General of Graduate Schools of Business and GovernmentITESM Ð Tecnol—gico de MonterreyDr. Hans-Martin FŸsselProject Manager for Climate Impacts, Vulnerability, and AdaptationEuropean Environment AgencyDr. Richard MossSenior Staff Scientist Pacific Northwest National Laboratory Joint Global Change Research Institute Dr. Mark MyersVice Chancellor of ResearchUniversity of AlaskaDr. Anand PatwardhanProfessor, Shailesh J Mehta School of ManagementIndian Institute of Technology-BombayDr. Caroline SullivanAssociate ProfessorSchool of Environmental Science and ManagementSouthern Cross University, AustraliaDr. Claudio SzlafszteinProfessor, Center of Environmental SciencesFederal University of Par‡, BrazilDr. Jintao XuProfessor of Natural Resource EconomicsChair, Department of Environmental ManagementNational School of Development, Peking University The Dimensions of the Global Adaptation Problems Required Resources$150-200 billion investments required for all adaptation needs per year Including É $14.4-19.7 to protect water supplies and against increased flooding, $27.6-28.5 to protect coastal zones and $13-27.5 to build and rebuild more resilient infrastructure Currently Devoted Resources$9 billion total global funding for climate adaptation AND mitigation$1.8 billion for adaptation through global funds Major Target for Filling the Funding Gap -- Private Sector Opportunities$100 billion + opportunities for annual investments Executive SummaryThe Global Adaptation Institute The Global Adaptation Institute is an international non-profit organization dedicated to assuring sustainable development by improving resilience against climate change and other natural disasters.  Our mission is to enhance the worldÕs understanding of the urgency for adaptation and to direct private and public investments to where they can be most productive. What makes us unique?¥	WeÕre not a ÒthinkÓ tank, but a ÒdoÓ tank.  Our goal is not simply to describe the problem of adaptation. We want to help fix it.  We go one step further and provide an exciting new tool to reveal where invested resources can create the greatest good.¥	We focus on the private sectorÕs role in adaptation in addition to that of governments and NGOs.   There is a multi-billion dollar gap between current government and NGO investment in adaptation, and what is ultimately needed.  The private sector could be a major source to fill that gap.The world is changing. Climate change is exacerbating social and environmental stresses: severe droughts in the Horn of Africa, unusually extensive flooding in Pakistan, Australia and the Midwest United States, extreme earthquakes in Haiti, Chile  and Japan followed in some cases by devastating tsunamis, and more tornadoes were recorded in the United States in April 2011 than in any previous month on record. More and more people are forced to deal with unexpected climatic and other natural disasters. The effects of climate change are compounding global trends such as population growth, migration and urbanization, and economic development. The Institute recognizes the importance of global climate mitigation policies and agreements. However, the climate will continue changing throughout this century whether or not a binding international climate policy develops. Therefore, it is incumbent upon leaders in government, industry, and all forms of civil society to prepare for both anticipated and unforeseen risks to human life and livelihood. In other words, we must take action.  We must adapt, and adapt in a way that is pragmatic, realistic and based on the principle that individuals should be able to take destiny into their own hands. Scientific InputWe are bringing together science, business and government leaders to create accurate metrics on adaptation.  The metrics must be pragmatic and useful to users. These metrics are summarized in the Global Adaptation IndexTM (GaInTM).  In addition we are researching complimentary demonstration projects and organizing outreach activities designed to move beyond just a description of climate vulnerability and to show how concrete actions can improve human life and increase the ability of nations to protect their citizens. Not Just a Private Sector FocusThe Institute aims to influence governments, investors, business leaders, NGOs and other adaptation enabling actors. Our GaInTM is designed to provide direct guidance on the best use of government, non-profit or donor resources. The need for greater direction is recognized globally.  We aim to provide that key tool. In the process we hope to identify the largest and most dynamic innovations for unleashing the human potential to adapt and to provide the most positive impact on human lives.  GaInª in Practice(A few summary examples)?The Global Adaptation Problem The climate is changing. Global temperatures increased by approximately 0.7¡C in the 20th century and are likely to rise by at least another 2¡C this century. Even drastic reductions in global anthropogenic CO2 emissions are unlikely to significantly abate this trend. These temperature increases will compound the current effects population growth, urbanization and economic development are having on humanityÕs ability to access and produce goods and services from natural resources. Water Ð Stable water resources underpin the location of cities, the ability of small businesses and multi-national corporations to profit, and the basic functioning of human households. Increased flooding, droughts and rainfall patterns have repercussions for human health, agricultural productivity, human migration and the operation of industry. Agriculture & Food - Billions around the world depend on agriculture for their livelihood, and hundreds of millions depend directly on their farm land for subsistence. Climate change may improve crop productivity in one place, yet decrease it in another.  Increasing variability in rainfall, floods, and seasonal temperatures across the globe threaten all farmers requiring alteration of agriculture management practices. Some farmers may not be able to grow their current crops or produce any food at all without migrating elsewhere.Energy - Energy plays a fundamental role in the economic, social and technological development of countries. People need energy to enhance their potential. Jobs are created when reliable energy is available. Health and education improve with access to energy. Fluctuations in water supplies complicate the operation of many fossil fuel-powered energy systems, while an increase in natural disasters put at risk infrastructure that provides energy to people. Coastal protection - Rising sea levels create challenges for any coastal nation, especially for people living near the sea. Land loss leads not only to a decline in the value of coastal property, but also to elimination of livelihoods dependent on fishing, farming and tourism.Transport Ð Transportation is not only crucial for commerce, but vital for the provision of basic goods and services necessary to protect lives and livelihoods. Floods can severaly damage all types of transportation with unpaved roads the most sucseptible to many weather events.  sanitation systems, electricity lines, and water lines are all affected by changes in temperature, the frequency of natural disasters, and water levels. The provision of basic services can be disrupted, restricting the ability of economies to grow. Health Ð human health is impacted by all sectors listed above. Even where these sectors remain stable, increases in heat waves and expansion of disease can take a toll on human life. ?The Global Adaptation IndexTMAll countries are struggling with the challenges facing these sectors. However, some are more vulnerable than others due to the location of their populations or geographic and topographic conditions. Further, some nations are more ready to deal with these challenges through government action, community awareness or the ability to facilitate private sector responses. The Global Adaptation Indexª measures all major aspects of a countryÕs vulnerability versus existing institutions or investments that either reduce the impact or make a country more ready to respond to specific disasters. This comparison of vulnerability versus readiness facilitates the distinction between countries in which the government or NGOs should invest or make institutional changes, and those most attractive to private investment. It can also point out countries in which little investment by either source is required. [Check for accuracy when GaInTM results are complete.]For example, where the convoluted path to obtaining an operating license is prolonged, securing financing is onerous or protecting basic assets and property is uncertain, the economic climate is not hospitable for doing business. Fewer businesses form, less capital is invested, and business cannot adapt as easily to changing conditions.  Less private sector effort will go into investments to protect, warn or provide relief to the people.  Unless the government has taken up the slack, when a disaster hits the damage is greater and the response smaller.  The suffering of people is more intense.Governments must make a choice.  Should they adjust their regulatory environment to encourage more private sector preparation and response, or are they prepared to make the sizable investments the private sector potentially could provide?  If they want the private sector to take on more responsibility, what must change?  The Global Adaptation IndexTM is designed to help sort out these adaptation choices. ?Vulnerability and Readiness: The Readiness MatrixThe Global Adaptation InstituteÕs Readiness Matrix presents the comparative vulnerability and readiness of countries. The ÒReadiness MatrixÓ (shown below) is an open canvass waiting for a scientific approach to determine where individual countries lie.  The vertical axis measures the relative vulnerability of countries.  The horizontal axis measures the degree to which a country is prepared to deal with climatic, socioeconomic and environmental changes.  The colored quadrants illustrate the meanings of the categories into which a country might fall:1.	Red or Upper Left Quadrant Ð A country with a high exposure to climatic change, but a low level of readiness, has both a great need for investment and innovations to improve readiness and a great urgency for action.  Unless the government, international organizations and the private sector move quickly to improve the ability to adapt, the country is unlikely to deal successfully with change.  Initially this country is more likely to receive investment from the government or NGOs than from the private sector looking for profitable returns.2.	Yellow or Lower Left Quadrant Ð By contrast, a country that has both low vulnerability and low ability to adapt faces few immediate challenges and therefore has more time to improve the ability to cope.  There are likely specific, profitable opportunities for the government, international organizations and the private sector to intervene.3.	Blue or Lower Right Quadrant Ð In countries with low vulnerability but a high ability to adapt, adaptation challenges may still exist, but these countries require the least help.  They are comparatively well along the road to responding effectively to sudden climatic or environmental changes.4.	Green or Upper Right Quadrant Ð This country is also on the road to responding effectively (high level of readiness), but the remaining needs are greater, as is the urgency to act (high vulnerability).  Because the country has shown an ability to raise its level of readiness, the private sector is more likely (than in the red quadrant) to participate in the effort to improve adaptability.  The Scientific Approach: The Global Adaptation Indexª Ð GaInªMeasuring 170 countries against 24 Vulnerability and 14 Readiness indicators GaInTM estimates the relative position of each country within the Matrix. The Vulnerability Axis captures a countryÕs exposure, sensitivity and capacity to adapt to the effects of climate change and other global trends on water, food & agriculture, health, coastal protection, transportation and energy, sectors important to human well being.  The Readiness Axis captures the economic environment, governance, and societal capabilities.  Each of these is important for determining how quickly and effectively countries can respond to sudden disasters.  Results from GaInTM 1.0(?) are illustrated in Figure 2.  Full ReportIntroductionThe world is changing fast. Countries are being challenged to prepare for and, if possible, to minimize the effects of climate change. The challenge is only greater as populations, urbanization, and the economy grow. Despite expanding resource commitments from international institutions, public funding alone is not the solution. The private sector must also play a key role in these investments. With appropriate information all can contribute to increasing the resilience of local communities. The Global Adaptation Indexª (GaInª) is a navigation chart that can guide investments both private and public. The Global Adaptation Institute www.globalai.org, a non-profit environmental organization, is developing GaInª to compliment our future adaptation pilot projects and global outreach efforts. We are guided by a vision of building resilience against climate change and other global forces as a key component of sustainable development. The Institute recognizes that mitigation continues to remain an essential global policy goal. However, the climate will continue changing throughout this century whether or not a binding international climate policy develops.  As history has shown, increases in climate-related disasters and climate change will lead to increased risks and costs for businesses, complicate political decisions, and of most concern, threaten the quality of life for vulnerable populations around the world. Therefore, it is incumbent upon leaders in government, industry, and all forms of civil society to prepare for both anticipated and unforeseen risks to human life and livelihood. In other words, we must adapt, and adapt in a way that is pragmatic, realistic and based on the principle that individuals should be able to take destiny into their own hands. What makes us unique?WeÕre not a ÒthinkÓ tank, but a ÒdoÓ tank.  Our goal is not simply to describe the problem of adaptation. We want to help fix it.  We go one step further and provide an exciting, new tool to reveal where invested resources can create the greatest good.We focus on the private sectorÕs role in adaptation in addition to that of governments and NGOs.   There is a multi-billion dollar gap between current government and NGO investment in adaptation, and what is ultimately needed.  The private sector could be a major source to fill that gap.What makes GaInª unique?GaInª is certainly not the first attempt at measuring the vulnerabilities of countries to climate change and other global trends. Scientists and institutions around the world have conducted significant work throughout the last decade, many listed in the Reference section of this report. GaInª seeks to build upon this previous work by creating an index that promotes pragmatic action among governments, the private sector and NGOs. Further, the index:¥	Is open and transparent Ð All indicators and sources are readily available and easily accessible by the public. ¥	Brings the private sector to the table - Most indices focus solely on the vulnerabilities of countries; GaInª includes indicators that guide governments and communities in how to harness the power of the private sector. ¥	Focuses on sectors crucial to human well being Ð GaInª does not cover all aspects of vulnerability, but targets those sectors most important to human health and prosperity that also can be greatly improved by innovation and action from the private sector. [Ian Is this statement correct?]¥	Provides estimates of the additional population protected from an improvement in vulnerability [True(?)].¥	É.Audience & UseThe GaInª model is intended to be a tool that decision makers in both the private and public sectors can use to save and improve the quality of lives. While there are tools that can assist policymakers, businesspeople, scientists, and other users in determining risks from climate change and other major trends, our goal is to create a facility that encourages users to move beyond descriptions and act. Our tool is intended to be pragmatic, actionable, and oriented toward delivering improvements in climate resilience through the enhancement of the social, governance and business environments in all countries. We believe that currently government institutions, international non-profits, and major development funders are further along in developing metrics that represent their priorities for adaptation than is the private sector; guidance that will mobilize action from the private sector has been scarce. Thus, the indicators mentioned above and described in more detail throughout this document have been selected to encourage private sector participation.  Effective corporations and investors observe and respond to issues important to government and civil society. Effective governments and NGOs work with the private sector. Thus, our Index should be useful for all actors in gauging where private capital can Òmove the adaptation needleÓ in a positive direction versus where NGOs, development institutions and governments need to step in.  The results from GaInTM will not have the same meaning to everyone. A country may appear inhospitable for business investment due to a low degree of readiness or high degree of vulnerability. However, that same country may be an appropriate target for development support from government and international institutions. Similarly the private sector may support programs that governments and international institutions deem marginal. For many in the private sector this is the tool to help decide where to invest both to obtain an attractive rate of return and help people in need.Limitations Coverage - Not every country has the requisite data, so not all countries are covered. Currently we consider 170 countries.  As data become available or more comprehensive, additional countries may be included in subsequent versions of GaInTM. Ideal indicators - Indicators have been chosen in part by the number of countries covered. Several useful indicators have been omitted because effective reporting is only available in industrialized nations or specific regions. The Institute is open to suggestions for additional indicators. Completeness of Readiness Ð The Index does not cover every indicator that represents whether a country is prepared to deal with climate change and other global forces. Demand - Is there enough awareness in a country or region on specific vulnerabilities to create a demand for services? Even if a country has a high degree of readiness, gaps in information may exist and media coverage or political will may be lacking. The Institute will attempt to address these questions in the future through the Adaptometerª project, describe in page ÉGovernment initiative Ð While a hospitable investment climate typically requires the presence of a responsive, effective and transparent government, there are exceptions in which a country may be an unnattractive destination for investment but have adaptation policies that save lives. While such countries may have useful plans in place, we believe even more progress could be made by harnessing the power of the private sector. Future work by the Institute and others can help paint a more complete picture of Readiness. For now, given the urgent need for private sector action, we believe an Index targeted to the private sector is most needed now. 									?Adaptation OpportunitiesRole of the Private SectorRisk mitigation Ð Insurance companies have been at the forefront of recognizing and quantifying climate change risks. Such risks can have a profound impact on an insurance companyÕs activities by either raising insurance rates or refusing coverage to all companies and/or projects that seem risky in light of anticipated climate change impacts. Other companies take unilateral steps to protect themselves. Many are highly dependent on natural resources, and some of these companies have also acted to improve the long-term quantity and quality of these resources. Coca-Cola, for example, has been actively engaged with local communities in watershed management. Pfizer aims to improve access to clean water as one of their strategic foci to mitigate production risks and improve quality of life at the same time. New products & services Ð Some companies are finding opportunities providing new products or services that will assess how communities and individuals cope with a changing climate. For example, communications companies such as Nokia have been providing farmers in developing countries with weather data delivered to their mobile phones. Insurance companies such as Allianz have offered additional affordable agricultural insurance in developing countries. Other companies have created subsidiaries to address specific needs. For instance, Tata created Tata Swatch to develop water purification products. Corporate Social Responsibility (CSR) Ð CSR is a leading driver of private sector involvement in climate adaptation. Companies that previously had a strong commitment to sustainability and humanitarian pursuits are now also supporting adaptation activities. For example, Honda is supporting marine protected areas by providing equipment, developing management plans, and public education efforts. Additionally, GlaxoSmithKline has reduced several of its pharmaceutical product prices up to 25 percent in least developed countries and reinvests 20 percent of profit-share in healthcare projects. Deutsche Post runs disaster preparedness training programs for its airport staff in several developing countries including Bangladesh, Nepal and India. Markets for ecosystem/adaptation services Ð companies that own large tracts of land can manage their holdings to enhance or preserve ecosystems that provide protection and resilience against climate change and other forces. Timber and agricultural lands can be managed to preserve waterflows and quality.  Role of Government and NGOsBoth governments and NGOs can help people adapt. The coordination of multi-state or regional adaptation projects will likely require government support. Likewise, many small-scale, local projects may need NGO and civil society buy-in. Further, environmental data, population and economic statistics, and other information relevant to the success of private-led adaptation solutions are often gathered and maintained by government institutions.Some urgently needed adaptation projects may not provide immediate profits or incentives for private sector involvement. However, they may still provide substantial benefits to the community as a whole.  Such situations may require government and community-based leadership. Governments, NGOs and other non-private sector institutions should find GaInª useful, as it will point to where gaps in adaptation funding can be filled by the private sector. Where they cannot, action by governments and other institutions may be needed. ?GaInª Structure	Vulnerability and Readiness: The Readiness MatrixThe Global Adaptation InstituteÕs Readiness Matrix summarizes the comparative vulnerability and readiness of countries. The blank ÒReadiness MatrixÓ (shown below) is an open canvass waiting for a scientific approach to determine where individual countries lie.  The vertical axis measures the relative vulnerability of countries.  The horizontal axis measures the degree to which a country is prepared to deal with climatic and environmental changes.  Indicators selected for The Global Adaptation IndexTM attempt to capture the concepts underlying the two axes. Vulnerability Axis: This axis seeks to capture exposure to climate related hazards, sensitivity to their impacts and the ability to cope with those impacts. Vulnerability has 3 categories representing the core components of vulnerability (exposure, sensitivity, capacity), and 6 key sector indicators for each category. For three sectors (water, food, health) two types of measures are incorporated. One describes essentially quantitative measures (amount, how much is used, how many have access) and the other quality measures (exposure to poor quality, impacts on people, access to quality management measures).  For coastal protection, infrastructure and energy, only quantitative indicators are used. Readiness Axis: The readiness axis measures how well countries can respond to slow challenges (impact of changing temperatures on crops, growing water shortages) and sudden natural disasters (floods, tornados, earthquakes).  Response is in part the speed and ability of a countryÕs private and public sectors to absorb additional investment resources and apply them effectively. It is also the ease and speed of getting these resources to affected areas. The axis has 3 categories: Economy, Social, and Governance. Selection of Indicators: Based on consultations and feedback, it was agreed that the vulnerability measures included in the GAINTM should not only fit within the above structure but also:1.	Meet user needs and be consistent with current knowledge and best practice.2.	Be transparent and conceptually clear3.	Be based on data that are freely accessible, quality checked, and comprehensive in national coverage.4.	Be potentially scalable from national to regional and local.5.	Avoid directly incorporate broad socio-economic measures, such as GDP/capita, HDI etc. .6.	Consist of existing relevant indices.  Where indicators have overlapping information (highly correlated), only one should be used [Correct?]. 7.	Include as many (UN) countries as possible.8.	Point to actionable improvements. Specifically for vulnerability, the Sensitivity and Capacity measures should be actionable; i.e. indicate actions that lead to measurable changes within a few years.9.	Have time series data available wherever possible so that national progress can be tracked.  ?The Vulnerability AxisVulnerable SectorsThe Vulnerability Axis of GaInª provides indicators for 6 sectors important to human well-being. Through appropriate incentives, the private sector can improve the provision of water, food, energy and other services and products. WaterReliable water supplies are the basis for economies and civilizations. Industry, agriculture, and human health can be immediately impacted by alterations in water levels and precipitation. Current population and economic pressures on water supplies will be compounded in coming decades by climate change. This will be particularly true at the regional and local levels as there is greater variability in the sustainability of water supplies within nations. Technologies that make water use more efficient will be in high demand. Companies that develop reuse and conservation processes for both industry and human consumption will profit. Lives will be saved through resilient water distribution systems. Industries, particularly beverage, energy and agricultural companies that dependent directly upon water supplies, will need to assess their operational and supply chain risks. Companies must increasingly work with governmental entities and stakeholders to protect and conserve water supplies, through cooperative agreements, water trading and pricing, legislation, or other measures that will ensure future provision of water resources. Food & AgricultureBillions around the world depend on agriculture for food or their livelihood, while hundreds of millions depend directly on the land they farm for their subsistence. Population growth, environmental degradation, and climate change all stress current agricultural production. While climate change may improve crop productivity in some places, increasing variability in rainfall, floods, and seasonal temperatures across the globe will cause all farmers to alter their agriculture management practices to at least some degree. At worst, farmers may not be able to grow their current crops or produce any food at all. Through research and development, the private sector can help develop more drought-resistant crops, irrigation techniques, and delivery systems. New insurance mechanisms may also help provide safety nets for farmers in the face of increasing uncertainty. HealthFor the millions who are not lifted out of poverty or who continue subsistence lifestyles during this century, small changes in climate may have significant effects. An increase in number and severity of natural disasters, whether flooding, hurricanes or droughts, can injure of kill those in communities not prepared for such events. Even in wealthy nations, extremes of heat and cold weather can inflict harm on many. Further, diseases such as malaria may spread as mosquitos and other parasites expand into warming regions. The private sector can help in a variety of ways, from investing in treatments for diseases expected to increase to bolstering traditional healthcare measures such as improving healthcare infrastructure and delivery systems. Moving to secure the health of communities and a healthy workforce will be essential for businesses both large and small.Coastal ProtectionRising sea levels will create challenges to any anyone living along coastlines or near the sea. Land loss leads not only to a decline in the value of coastal property, but elimination of livelihoods dependent on fishing and tourism in these areas.Storm surges from open water will have a greater impact on low-lying cities and communities and endanger the fresh water supplies of millions living near the coast. Small business need to prepare for possible damage to infrastructure (e.g. roads and electricity lines) vital to their operations. While large corporations face the same challenges , they are in a stronger position to invest in resilient strategies, whether infrastructure (sea walls) or in ecosystem services like mangrove protection. Opportunities will continue to emerge for the private sector as municipalities increasingly need assistance facing freshwater supply, sewage, population relocation and transport challenges. Energy, trade, and agricultural systems will be affected as well, requiring new thinking and problem solving from industry. EnergyEnergy plays a fundamental role in the economic, social and technological development of countries. People need energy to enhance their potential. Jobs are more easily created where reliable energy is available. Health and education improve with access to energy. Major global trends, including climate change affect all energy systems, from traditional wood burning to industrialized fossil fuel and renewable energy systems.The availability of water underpins threats to energy supplies. Coal-fired and nuclear energy utilities depend on large quantities of freshwater for cooling. Investors in hydroelectric power need to plan for greater variability in water flows.Increased natural disasters can disrupt energy transport and transmission routes as well as physically damage energy facilities. Changing regional temperatures, rainfall patterns and ranges of invasive pests may also complicate harvesting of biofuel crops.Inappropriate regulation impedes investments in energy.Transport Poor transportation infrastructure can impede not only economic development, but the ability of nations and communities to save lives through the delivery of medical services and food supplies. The private sector can build roads É?Vulnerability Components & Indicators The vulnerability component of GAINTM follows a consistent structure based on the best of published thinking on the topic, the IPCC Review process, and guidance received from the Advisery Board and Council of Scientific Advisers of the Global Adaptation Institute.  The Vulnerability Axis of GAINTM seeks to represent vulnerability and adaptive capacity that is largely in the domain of action by communities, Civil Society Organizations and/or national governments.  Adaptive capacities that may realistically be influenced by or affect the private sector are better reflected in the Readiness Axis.The basic structure focuses on an initial set of ÔsectorsÕ made up of Water, Food, and Health that cover core issues relating to peopleÕs well-being.   It also has an infrastructure section that initially includes the three sectors; Coastal, Transport and Energy, which directly affect the livelihoods of many people.  Within each of the sectors vulnerability is represented by three components describing Exposure of the sector to climate related hazards and in particular to climate change; Sensitivity of that sector to the impacts of that hazard and the Capacity of national populations to cope/adapt with these potential impacts.  The vulnerability index is completed by selecting a series of quantitative measures to represent each element of the structure.  The effectiveness of the vulnerability depends on the quality of these choices. The GaInTM 1.0 model implements this guidance as follows.¥	For each of the sectors (water, food, health) two streams of measures are incorporated; one describing essentially quantitative measures (amount of water, how much is used, how many have access) and the other, where possible, quality measures (exposure to poor quality water, impacts of poor water on people, access to quality management measures).  ¥	The selection of measures will always be a subjective exercise, but some greater objectivity is introduced by following the above guidelines and the systematic structure.  Numerous variables (about 20 different measures for water alone) were considered and tested against the guidelines in making the selection.  ¥	Often potential variables were highly correlated.  In this case they were examined for the information that appeared to be associated with the ÔoutlierÕ observations to see if the variables had different sensitivities.  They were also compared with other measures to be included in a sector, or other sectors, to see if they appeared to be providing additional information.  Often the choice was swayed by the availability and quality of the global data.	The Table below briefly describes the selection for GaInTM 1.0.  They are subject to revision based on further analysis and advice from the Council of Scientists, the Advisory Board and wider public consultation.  Technical Annex 1 provides a more detailed technical description of each of the measures and the data sources used.  The highly structured inclusion rules (sectors, by components of vulnerability etc.) provide a basis for a debate on the removal and replacement by better measures, inclusion of new sectors etc.Vulnerability Indicators OverviewSector	Exposure	Sensitivity	CapacityWater	Quantity	Projected change in precipitation	Water extraction	Water access	Quality	Projected change in temperature	Water-borne disease mortality	Water sanitationFood	Quantity	Projected change in agricultural yield	Rural population percentage	Agricultural technology  	Quality	Calorie deficit	Food (cereal) imports	Child nourishmentHealth	Quantity	Frequency of disease	Health professionals numbers	Life expectancy	Quality	Deaths from disease	Health expenditure 	Maternal mortality Coastal	Quantity	Land elevation	Population in low elevation	Measured on the Readiness AxisInfrastructure	Quantity	Infrastructure	Reliable roads	Energy 	Quantity	Access to reliable electricity 	Energy use	Methodology Systematic methods for converting the raw data from the various measures into an index are needed.  There are many different methods for rescaling raw data into similar ranges of values, including normalizing to a common mean and standard deviation; setting base low and high values for the raw data (e.g. from the observed minimum to the observed maximum; or from 0 to 100% compliance etc.) and scaling the data either linearly or after transformation to a prescribed range (e.g. 0 to 1; 0 to 100; -1 to +1).  Some indices convert the raw data to ranked values and work with these rankings.In the GaInTM Vulnerability Index, we were guided by the principles of transparency and simplicity.  Thus, we used the following procedures. ¥	The measure was selected so that the raw data itself was a meaningful representation of the information we were seeking.  If some form of transformation was needed (e.g. expressing in appropriate units, log transformation to better represent the real sensitivity of the measure etc.) it was considered at this stage.¥	Baseline minimum and maximum values for the raw data were selected.  These were based on common sense criteria and usually encompassed all or most of the observed range of values across countries, but in some cases the information in the measure saturates.  For example, for access to improved water the range was set from 0% to 100% as any improvement in access is likely to reduce vulnerability, but for mortality from communicable diseases it was set from 0 % to 50% even though some countries have much higher rates.  This gives greater sensitivity over the range in which most countries fall and greater incentive to improve health services.¥	All measures were then scaled to the range 0 to 1.¥	The measures were combined by simple arithmetic average for each of the sectors considered without further weighting.  In GAINTM 1.0 this means that each of the four sectors is the average of six measures.¥	The sectors were combined to give the final Vulnerability Index by simple arithmetic average again.Other methods for calculating the final index are possible and could be justified, but this is a simple and transparent process and it appears to provide an adequate result.  Technical Annex 3 provides a discussion of the sensitivity of the Vulnerability Index to these assumptions including the use of geometric means, ranked values and the random exclusion of measuresData Availability The availability of globally comprehensive and quality checked data at national level data is one of the main criteria for selecting measures for inclusion in the GaInTM.  A number of measures, including some often used or cited as good indicators, such as measures of irrigation use, were rejected because of poor or inconsistent coverage.  Data were taken from the World Development Index (May 2011 version) where available as this source represents a set of quality checked, maintained, full sourced and publicly available information.Table 2.  The 20 least and the 20 most vulnerable countries  and a sample from the middle ranked countries based on the 185 countries with sufficient data.Vulnerability	 Least	Middle	Most	Not ScoredDenmark	0.102	Tunisia	0.290	Lesotho	0.496	AndorraNorway	0.125	Indonesia	0.293	Marshall Is	0.497	Congo (DRC)France	0.139	Montenegro	0.293	Kiribati	0.500	LiechtensteinGermany	0.140	Saudi Arabia	0.293	Rwanda	0.505	MonacoUK	0.142	Albania	0.295	Tanzania	0.505	NauruFinland	0.144	Vietnam	0.295	Malawi	0.507	RomaniaSwitzerland	0.144	Libya	0.297	Afghanistan	0.507	Timor-LesteCzech Rep	0.145	Bahrain	0.299	Burkina Faso	0.511	TuvaluNew Zealand	0.147	Grenada	0.301	Mauritania	0.512	 San Marino	0.153	Fiji	0.303	Liberia	0.512	 Poland	0.158	Suriname	0.304	Sierra Leone	0.523	 Sweden	0.158	Sri Lanka	0.304	Mali	0.528	 Ireland	0.159	Bahamas	0.306	Chad	0.533	 Austria	0.165	Seychelles	0.307	Ethiopia	0.534	 Luxembourg	0.168	Philippines	0.310	Angola	0.535	 USA	0.176	Cape Verde	0.311	Zambia	0.546	 Uruguay	0.182	Nicaragua	0.312	Niger	0.558	 Slovenia	0.187	Costa Rica	0.313	Mozambique	0.560	 Italy	0.189	Dominican Rep	0.313	Cent Afr Rep	0.575	 Croatia	0.189	Sao Tome Princ.	0.313	Burundi	0.582	 The target set of countries for inclusion is the 192 members of the UN (as of 30 June 2011).  This means that territories etc. are omitted, but often data coverage for them is spotty.  Eventually full data sets were available for 171 countries and if the coverage criteria is relaxed to allow for up to 3 missing measures the number increased to 185.  Among the countries with missing data are several developed countries that do not report data consistently.  These include Andorra, Liechtenstein, Monaco, Romania and San Marino.  Sometimes data for the new Baltic States had to be estimated from earlier national statistics.  Data for Least Developed Countries and Small Island States were often missing.  Estimates were made where they could be logically justified, for example by assessing Small Island States as being fully exposed to coastal risks.  In some cases estimates were made that were moderately more pessimistic that the likely value.  This provides an incentive to the country to provide a more accurate estimate.  These estimates are documented in the full databases.Table 3. Sample of rankings based on the 185 countries meeting the ÒrelaxedÓ criteria for completenessVulnerabilityMost exposed	Most sensitive	Least capacityZambia	0.564	Burundi	0.587	Somalia	0.858Central Afr. Rep	0.553	Afghanistan	0.555	Chad	0.855Afghanistan	0.502	Somalia	0.546	Niger	0.848Mozambique	0.494	Marshall Islands	0.539	Sierra Leone	0.826Mali	0.494	Malawi	0.536	Afghanistan	0.825Bangladesh	0.490	Mozambique	0.535	Mali	0.796Botswana	0.488	Yemen	0.519	Nigeria	0.784Gambia	0.477	Kiribati	0.504	Ethiopia	0.783Lesotho	0.468	Ethiopia	0.502	Burkina Faso	0.773Angola	0.465	Papua New Guin.	0.502	Madagascar	0.751Burundi	0.465	Niger	0.496	Central Afr. Rep	0.751Dominica	0.464	Micronesia, F.S.	0.494	Burundi	0.749Mauritania	0.462	Samoa	0.494	Angola	0.722Togo	0.460	Senegal	0.493	Sudan	0.719Zimbabwe	0.456	Zambia	0.489	Guinea-Bissau	0.719Rwanda	0.455	Lesotho	0.488	Liberia	0.711Malawi	0.448	Rwanda	0.487	Guinea	0.707Namibia	0.447	Central Afr. Rep	0.481	Mozambique	0.696Myanmar	0.441	Angola	0.480	Mauritania	0.689Tuvalu	0.435	Bhutan	0.477	Tanzania	0.686What the Vulnerability Axis Tells UsThe full ranking is provided in Annex 2 and a summary in the Table 2.  The least vulnerable are a group of developed and largely European countries, while the most vulnerable are a mix of least developed African and Small Island States.  Table 3 shows the states that scored highest as the most exposed, sensitive or lowest capacity.  Again African and Small Island States predominate.  Four countries rank in the most vulnerable for all three components: Afghanistan, Burundi, Central African Republic and Mozambique.  Somalia appears high on both sensitivity (3rd) and low capacity (1st), but is not ranked in the top 20 most vulnerable states because it is assessed as having only moderate exposure.Relationship with Other Vulnerability IndicesA number of indices of vulnerability have been developed or proposed, but none has found widespread use.  This arises partly because vulnerability is very much affected by local factors, both environmental and social-economic.  Thus indices based on national level data are seen of little direct value.  Nevertheless national level indices have been developed often with the goal of prioritizing the allocation of resources, including finance.   We compared the GaInª Vulnerability Index with a selection of published indices (see also Box 1 for comparisons with sectoral indices such as the Water Poverty Index).  While there is strong agreement among the indices in that they find a group of least developed or conflict affected states as the most vulnerable, they vary significantly in the detailed rankings.  Some of this variation arises because of different goals for the indices and thus the selection of measures; for example, some are more oriented towards environmental quality, some are tuned for financial allocations etc.  Some of the variation arises from different approaches to the construction of the index (e.g. Brooks & Adger 2003); and some arises because the indices in the comparison have been developed over more than a decade and the available data and the condition of the countries have changed over that period.  Five indices and the GaInTM Vulnerability Index were compared by looking for overlap between countries in the top 20 for each index (Table 4).  A total of 65 countries appear in the 120 slots in the comparison (6 indices by 20 top ranks); 17 appear in 3 or 4 indices and none in more than 4.  The GaInTM Vulnerability Index includes 14 of the 17 among its top 20, which shows that it is not an outlier among the indices in its ranking of most vulnerable counries.  A closer comparison, however, shows differences in detail.  The Wheeler 2011 index has been developed at about the same time as that as GaInTM and with similar although not identical goals; Wheeler suggests his index as a tool for the allocation of public money for adaptation while the GaInTM seeks to provide guidance to both the private and public sectors on where to place investments.  The two indices have 11 countries in common among their top 20 but Fig. 2 shows that the correlation in scores between the indices is only r2=25%.Table 4.  A comparison of the 20 most vulnerable countries in recently published indices.  Countries in blue appear 4 times in the table and those in red 3 times.Moss et al 2001	WEF 2002	Brooks & Adger 2003	DARA 2010	Wheeler 2011 	GAIN1.0  Vulnerability IndexBangladesh	Angola	Antigua & Bar.	Afghanistan	Bangladesh	AfghanistanCambodia	Burundi	Australia	Angola	Burundi	AngolaChina	Cambodia	Bangladesh	Burkina Faso	Cambodia	Burkina FasoEgypt	Central Afr Rep.	Belize	Burundi	Central Afr Rep	BurundiIndia	Chad	Cambodia	Congo DR	Chad	Central Afr RepIran	Ethiopia	China	Djibouti	Congo	ChadLibya	Guinea	Djibouti	Equ. Guinea	Ethiopia	EthiopiaMexico	Guinea-Bissau	Fiji	Guinea-Bissau	Guinea	Guinea-BissauNigeria	Haiti	Guyana	Haiti	Guinea-Bissau	KiribatiPoland	Liberia	Iran	Liberia	Lesotho	LesothoRep Korea	Madagascar	Kenya	Madagascar	Malawi	LiberiaSaudi Arabia	Malawi	Kiribati	Mali	Mali	MalawiSenegal	Mali	Laos	Mozambique	Myanmar	MaliSouth Africa	Mozambique	Malawi	Myanmar	Niger	MauritaniaSudan	Niger	Moldova	Niger	Rwanda	MozambiqueThailand	Rwanda	Mongolia	Rwanda	Senegal	NigerTunisia	Sierra Leone	Philippines	Senegal	Sudan	RwandaUkraine	Somalia	Swaziland	Sierra Leone	Vietnam	Sierra LeoneUzbekistan	Zaire	Tajikistan	Somalia	Zambia	TanzaniaYemen	Zambia	Zimbabwe	Vanuatu	Zimbabwe	Zambia3	15	3	10	13	14The above numbers are a count of the number of frequently appearing countries (3 or 4 times).Correlation with National Income Per CapitaMany developing countries and both public and NGO development agencies have expressed concerns about vulnerability indices as they so often incorporate and simply exaggerate the effect of low resources, and in particular the effect of low incomes.  In selection the measures for the Vulnerability Index direct and indirect inclusion of GDP per capita was avoided.  However, many of the measures are correlated with GDP per capita and as a consequence so is the Index itself (Table 5, Fig. 2).  This is particularly true of the Capacity component and for the Health sector, as might be expected.Figure 1.   A comparison of the most vulnerable countries in the Wheeler (2011) and GAINTM 1.0 Vulnerability indices.  Even though there is overlap in the countries ranking as the most vulnerable, there are still substantial differences in the scores. Figure 2.  Relationship between the GAINTM Vulnerability Index and national income The scatter about the best-fit line, which can be considered as a predictor of Vulnerability based on national income, contains useful information about national circumstances.  Countries falling well above the best-fit line tend to be resource rich counties where the distribution of the benefits of the wealth is often not evenly distributed and vulnerability remains high.  But countries with high (or low) exposure, which is only weakly correlated with income, will also tend to fall below (or above) the best-fit line.  A web-based ÒDashboardÓ that will allow users to explore the richness in the almost 100 measures that go to make up the GaInª1.0 is under development .?Table  5.  Pearson r2 correlation  (with sign) of the components of the Vulnerability Index with Ln(GDP(PPP) per capita)Summary	 		Total	-0.72 		Water	-0.37 		Food	-0.47 		Health	-0.71 		Infrastr	-0.36 		Exposure	-0.47 		Sensitivity	-0.50 		Capacity	-0.72 		Water	 	Food	  	 		 Delt PPt as %	NS	ACDI	-0.15 %  f/water withdrawals	NS	% Rural Popn	-0.47 Improved water	-0.52 	Two measure Av	-0.34 Delta Temp 	0.08 	CV annual Yield	NSWater Deaths <5yr	-0.48 	Import Depend	0.06 Improved sanitation 	-0.64 	%Wasting	-0.52  	 	 	 Health	 	Infrastructure	  	 		 CC induced Dalys	NS	Area % <5m asl 	NS2*Doc+Nurse/Midwife	-0.52 	Pop % <5m asl	NSLongevity	-0.64 	% pop with electricity	-0.56 % mort comm diseases	-0.60 	Energy at risk	NSExt Resources Health	-0.42 	Floods frequency	NSMaternal Mortality	-0.52 	% Paved roads	-0.38  	 	 	 Testing the Validity of the Vulnerability IndexThere is no direct way to test indices that seek to provide information about future conditions, and especially where those conditions are expected to change from recent and current conditions.  Nevertheless it might be expected that an effective index of future vulnerability largely based on measuring actions taken recently also should be related to current vulnerability.  That leaves the question of how to measure current vulnerability.One test is that numerous studies have concluded that vulnerability to climate and related impacts is associated with low national incomes per capita.  As shown above, even though we were careful to minimize the direct or indirect inclusion of GDP per capita information from the GaInTM , we found a high correlation between the two.Another test is to ask if the GAINTM Index is correlated with recent impacts of climate change on people.  Here we use the CRED database on disasters to calculate the number of people affected by climate related disasters over the 30-year period 1980 to 2010.  The correlation was highly significant although with an r2=22% only.  However, as Fig. 3 shows that the probability that an individual has been affected by a climate related disaster in the past decade increases from less than 1% in countries with low values of the Vulnerability Index to 35% at the highest values.Figure 3.  The relationship between the GAINTM Vulnerability Index and the probability of being affected by a climate related disaster over the past 30 years. ?Testing the Derivation of the IndexSeveral tests have been performed to see how sensitive the outcomes from the index are to the method by which the index was calculated.One test is to ask how sensitive the index is to the set of measures that were chosen for inclusion.  Would the results differ greatly if some measures were omitted and others included?  What is the effect of a few measures being missing for a country?  It is not possible to estimate the effect of measures not chosen, but we can ask whether the index would vary if certain measures had been omitted.  Figure 4. The results from a Monte Carlo simulation of omitting particular measures from the Vulnerability Index.  See text for an explanation. To test this we ran 100 ÔMonte Carlo simulationsÕ in which we recalculated the index but each time randomly omitting a third of the measures that constitute the Vulnerability Index and then recalculating the country ranks.   In Fig. 4 the countries are plotted in their original (all measures) rank order across the X-axis and the Y-axis shows in how many simulations each country was ranked among the most or least vulnerable 25 countries (symbols).  We also created a random set of data for the measures and did a similar Monte Carlo simulation with this random set and the results are show by the lines.  The results show that random omission of measures has only a small impact on the top and bottom rankings and especially on the most and least vulnerable scores.  For example the countries ranked at about 25th in vulnerability (top or bottom) still had 60% chance of being in the top or bottom 25 with the random omission of a third of the measures.  Those ranked in the top or bottom 10 have almost 100% chance of remaining in the top 25.  Comparison with the random data shows that this tendency was far greater that for random data, i.e. than due to chance.This suggests that the Vulnerability Index is robust to the inclusion or exclusion of the particular measures.  It also suggests that calculating the Vulnerability Index for countries with a few missing measures probably still gives a good estimate of their GaInTM score.There are other ways the Index could be calculated.  For example, sometimes a geometric mean  is used.  However, geometric means are only sensible where the model underlying the index is appropriate because it will score a country that has any one measure scored as 0 as having no vulnerability.  Our model for the GAINTM Vulnerability Index is not compatible with calculating geometric means as we have used what is essentially an underlying additive conceptual model.Another approach is to use rank scores for each of the measures rather than the absolute values and then to average these ranks to get an overall index.  This was tried for components of the GaInTM data.  It showed than on the whole the results were similar, but that the ranked approach becomes less sensitive at higher vulnerabilities.  This is because where there are a number of countries with similar scores, the additive approach used in GaInTM gives them similar index values, whereas the ranking method will spread the cluster of countries over a wide range of ranks leading to wide differences in the final index.  This is shown in Fig. 5, which shows a comparison of the ranking and the GaInTM additive method in the water sector.  The average difference in final ranked position of a country for the two different approaches is 20 positions.Figure 5.  A comparison of a ranking method to calculate a Water Sector index and the additive method used in GAINTM.?Readiness?Readiness Components & Indicators The Readiness Axis seeks to measure the ability of a countryÕs private and public sectors to absorb additional investment resources and apply them effectively towards increasing resiliency to climate change. This focus is similar in spirit to the World BankÕs Country Policy and Institutional Assessment (CPIA) measurements, which determine development assistance for low income countries. The assistance is based on performance in implementing policies that promote economic growth and poverty reduction.The readiness concept targets those portions of the economy, governance and society that affect the speed and efficiency of absorption and implementation.  It attempts to capture for each of these three components how well the prevailing environment facilitates adjustment to change.  Economy If the private sector can easily and quickly set up new businesses, access new capital, refocus as conditions change and obtain resources from other countries, that private sector will be more ready than the private sectors of other countries that continue to experience considerable hurdles to business creation, such as securing credit, changing the nature of products, or moving resources across the border.  Governance If resources are brought into a country with rampant bribery and a weak rule of law, chances are good that they will be less efficiently used than in a country where none are skimmed off the top and agreements are dependable.  Social If residents can readily plan how to organize the new resources and collaborate with people in other parts of the country or world, the outcome will be better than in countries where people have little experience organizing, do not understand the potential of the resources and have no way to consult with others elsewhere on implementation.Readiness also represents aspects of adaptive capacity that can be influenced relatively quickly.  If businesses do not face severe barriers, restrictions or monopolies, they can more easily adjust to the new demands on the size, scope or focus of their endeavors.  If the rule of law is lax, governments can make changes to enhance reliability.  Improving a farmerÕs warning of weather emergencies can sometimes be as simple as buying a mobile phone.Selecting the Readiness IndexUnlike for the vulnerability axis, there is a fairly short list of closely applicable candidates for Economic, Governance or Social Readiness.  The two chief candidates for the Economy component are The Doing Business Index (World Bank) and The Index of Economic Freedom (The Heritage Foundation/The Wall Street Journal), and its subcomponents.  Each is widely circulated and used.  There are also some lesser known indices like the Economic Freedom of the World (Fraser Institute/CATO Foundation).  Candidates for Governance readiness include subcomponents of the Index of Economic Freedom, of the World Governance Index (World Bank) and the Failed States Index (The Fund for Peace). Social readiness focuses on four important aspects: Human capital, mobile communications, the freedom of workers, and the rule of law. Human capital measures the availability of a highly educated workforce.  The greater the percentage of highly educated people, the more likely potential problems will be anticipated, prevention and/or emergency plans will be drawn up, and people will be mobilized in a timely fashion.  The more workers are free to move about the country or be hired by refocusing businesses, the easier and faster problems can be addressed.  The stronger the rule of law, the more respect for societyÕs rules and enforcements, and the more confidently and faster the response of those with needed resources.Mobile communications continue to develop applications that enable citizens to accomplish tasks more easily and conveniently (e.g. electronic transactions, mobile banking, finding destinations).  In emergencies mobile phones can assist in more effectively organizing warnings and relief efforts (enhancing coordination of efforts, providing health care in remote areas, quickly identifying locations most in need, enforcing transparency, creating and accessing valuable information).Selecting the various indices and subcomponents was based on these criteria:1.	A minimal set of indicators should produce a relevant score to measure whether the country will be able to successfully implement adaptation projects.2.	As many (UN) countries as possible should be included.3.	As many years as possible should be included, ideally 1995-2011.  A useful time series allows the construct of readiness changes, enhancing the goal of tracking the trajectory within countries.4.	Indicators should be as credible as possible.5.	Data should be as generally available as possible (preferably publicly available) so that others could reproduce the results.6.	Scores should point to actionable improvements.Table 1 illustrates part of the process.  This table provides the correlations (R-squares) among the overall and sub-indices of the Failed States Index (FSI), the World Governance Index (WGI), the Doing Business Index (DBI) and the Index of Economic Freedom (IEF), plus the percentage of population enrolled in tertiary education (regardless of age)  and the number of mobile phones per 100 people .  Each variable uses the most recent data available.Table 1The Correlations (R-Squares) Among Candidates for ReadinessImmediately it is clear that information contained in one indicator is often captured by others.  The Failed States Index for instance turns out to be very similar in their definitions and correlations to components of the World Governance Indicators (correlations up to 0.6-0.8) and components of the Index of Economic Freedom (0.5-0.8).Some of the Doing Business (DBI) components are not strongly correlated with other candidates, but the overall index is correlated with at least three components of the World Governance Index.  It is also highly correlated with the overall IEF (0.7) and its Business Freedom component (0.8). These high correlations are not surprising. DBI consists of many measures of ease of business as well as some approximating rule of law. An important drawback of DB is that only two years of overall scores are available (though there are raw scores for each of its subcomponents back to 1995).Limitations for the World Governance Indicators are also immediately apparent.  Four of the six components are highly correlated with each other (as high as 0.92).  Clearly there is overlap of information in these components, and inclusion of all is inappropriate. The two exceptions ("Voice and Accountability" plus "Political Stability and Non-Violence") are significantly correlated neither to other WGI components nor to the overall IEF.  These two exceptions can clearly add to the overall information.The Index of Economic Freedom closely follows the six criteria listed above.  It is one of the most complete in the number of years and countries and is readily available (both for data and methodology). Its 10 subcomponents span not only the economy component (employing some parts of Doing Business such as number of days to start a business), but also (as just mentioned) the government component corruption and the social component rule of law . No wonder that correlations with the IEF seem to capture much of what other variables target. It is also directly actionable. A country with a low readiness score might find that it could improve its score by making it easier to start a business, by reducing corruption or increasing the independence of judges, or by allowing a freer flow of capital from the domestic or foreign financial markets. In addition there is statistical evidence that changes in IEF scores relate to the growth rates of countries (Figure 1) .  Changes in IEF scores are computed over a 10 year period (2000Ð2010) and ranked from the country with the biggest improvement in score to the one with the biggest decline. The ranking is divided into quintiles.  Within each quintile the average growth rate of all countries (contained in that quintile) is computed.  Those averages are the numbers shown above each quintile bar in Figure 1.  Clearly, the average growth rate for countries that had the biggest improvement (4.2 percent) is significantly larger than in the remaining four quintiles (2.3-3.2 percent).   Also the average growth rates for the next two quintiles (3.1 and 3.2 percent) are larger than those in the bottom two (2.3 and 2.6 percent).  The implication is that those countries improving their scores the most tend to have higher growth rates.Figure 1 However, two of the IEF subcomponents (Property Rights and Freedom from Corruption) are highly correlated (0.9) to two from WGI (Rule of Law and Control of Corruption).  The source of the high correlation is not surprising since they are targeting the same aspect of governance and society.  Therefore the IEF subcomponents are freely interchangeable with the corresponding ones in WGI.The remaining two society indicators are far less correlated to other variables (Table 1).  For the number of mobile phones per 100 people, the correlations range from 0.00 (IEF Fiscal Freedom) to 0.56 (Failed States Deterioration of Public Services).  For the tertiary school enrollment the correlations range from 0.1 (IEF Labor Freedom) to 0.61 (Failed States Mounting Demographic Pressure). The final GaInTM version 1.0 of the Readiness Axis is captured with the combination consistent with Table 1 results.  Fourteen variables: ¥	Eight of the 10 subcomponents of the IEF¥	The two components from WGI that are uncorrelated with other indicators (Voice and Accountability and Political Stability and Non Violence)¥	The WGI Control of Corruption plus the Rule of Law indicators in place of the corresponding two from the IEF ¥	The percentage of population enrolled in tertiary education plus the number of mobiles phones per 100 residents.  It should be noted that while the tertiary school enrollment satisfies most of the criteria laid out, there is data for less than the average number of countries of most data series.  This data shortage complicates including a number of countries on the Readiness Axis. More extensive coverage of readiness is desired, but we have been unable to find as good a measure of advanced human capital that also has better coverage of years and countries.Subsequent versions may consider additional potential variables, and we encourage suggestions of other candidates.  Their relevance, however, will be determined in the same manner.Weighting the Fourteen Variables and Creating the Readiness AxisThe Readiness Axis is created through a relatively simple process.  The weightings are divided among the three components of readiness (economy, governance, society) and fourteen indicators as follows: The Economy component is assumed most important, so it is given the largest weight (40 percent).  The remaining two components are weighted equally (30 percent).Within Economy, the 8 relevant components from the Index of Economic Freedom are weighted equally (5.71 percent).  The three indicators for Governance also receive equal weight (10 percent).  The four Society indicators, however, are not equal.  It is assumed that education and the rule of law are essential for any countryÕs readiness, and these two receive twice the weight of mobile phones and labor freedom.The values of each of the fourteen variables are normalized between 0 and 1 by dividing each countryÕs number in the series by the maximum value found among the countries.  The scores are then appropriately weighted and aggregated.  Rankings are based on the aggregate score.Mobile phones and tertiary education have one additional step in the aggregate score process.  It is assumed that once the number of mobile phones per 100 reaches a certain level, that measure of Social Readiness is saturated.  In other words, while an additional phone per 100 people may add to the readiness when the number is 25, at some point enough phones exist that another makes little impact.  Clearly where there are 150 phones per 100 people, raising the number to 151 adds little to preparation.  It is assumed that the point of saturation is 90 (though it may well be below that level). All countries where the number of phones is 90 or above receive the maximum standardized value of 1.  A similar cap is applied to the proportion of population in tertiary education.  A high level of human capital increases the ability to effectively absorb resources, but as that level rises increasing less is gained from additional advanced degrees.   For this indicator the limit is assumed reached when 70 percent of the relevant age group is reported to be enrolled in tertiary education. Not surprisingly the aggregate rankings of readiness are highly correlated to the wealth of the country.  Table 2a and 2b illustrate this point for the top 10 and bottom 10 countries on the Readiness Axis.  The top-ranked countries are highly developed, and the bottom is dominated by some of the poorest of the poor countries.  Table 2aTop 10 Countries1	New Zealand	0.862	Australia	0.853	Denmark	0.844	Finland	0.835	Switzerland	0.836	Ireland	0.827	United States	0.818	Sweden	0.819	Canada	0.8110	Norway	0.81Table 2bBottom 10 Countries149	Nepal	0.37150	Cuba	0.35151	Yemen	0.35152	Comoros	0.35153	Burundi	0.34154	Ethiopia	0.32155	Chad	0.31156	Central African Republic	0.31157	Congo, Dem. Rep. of the	0.27158	Eritrea	0.24A scatter plot of the Readiness Scores and the log of Gross Domestic Income per capita (GDPpc) (Figure 2) confirm this relationship.  There is a very high correlation (0.8) between the readiness score and level of income per capita.  FIGURE 2Readiness Score Versus Income Per Capita However, the plot also reveals that there are important differences among countries even with similar levels of income.  If we consider the trend line as the global relation between readiness and income per capita, countries above the trend line have more readiness than their income levels would indicate.  These are the countries that may be attractive to private investors.  In contrast the countries below the trend line have less readiness than anticipated and are far less attractive for investment.In other words there is far more useful information in a ranking of readiness that is adjusted for GDP per capita.  Top ranked countries are likely no longer dominated by the richest countries.  Instead, they will be the ones who have created the most readiness for their size.  Consequently, the Readiness Axis is computed by subtracting for each country the trend level of readiness corresponding to the countryÕs GDP.Table 3a and Table 3b illustrate this point by showing the Top 20 and Bottom 20 countries using this trend-adjusted readiness ranking process.  Countries of different incomes are now scattered throughout the list.  For example,  Ghana, a lower-middle-income country (World Bank) is now ranked in the top 10 of readiness.  Mozambique and Madagascar, lower-income countries are in the top 20.  The United States, Canada, Sweden, and Norway no longer can be found in the top 20.At the other end of the scale, upper-middle-income economies the Russian Federation and China are in the bottom 20.  Low-income economies Nepal, Ethiopia, Eritrea, Burundi, Chad and the Central African Republic are no longer at the very bottom of the list.?Combining the Indices The vulnerability and readiness scores for each country are combined to produce two distinct results: a ranking and a matrix. To create a global ranking, the readiness score is subtracted by the vulnerability score. This combination is intended to show the overall preparedness of a country in protecting its citizens against the impacts of climate change and other global trends by using resources from the private sector. Such a ranking can clearly show how a country could move up or down the ranking by improving upon specific vulnerability and/or readiness indicators. The matrix (Readiness Matrix) is a tool most useful for the private sector to make specific investment decisions based on their desire for risk and adaptation needs. A country in the upper right quadrant (high vulnerability and high readiness) would most likely be targeted by conservative investors whereas those attracted to risk might seek out opportunities in the upper left quadrant (high vulnerability and low readiness.  To calculate a countryÕs position within the matrix, the vulnerability and readiness axes are weighted __ and __, respectively. While the ranking and matrix do not convey the same information, they are related by the same underlying indicators and movements in one tool will result in movement in the other. For instance, movement up the ranking due to improvements in vulnerability could move a country from the upper left quadrant to the lower left quadrant. The exact same ordinal movement along the rankings, instead, due to improvement in readiness, would move this country from the upper left quadrant to the upper right quadrant. Results in Detail ?Future WorkAdaptometerªRecognizing that many of the impacts of climate change and solutions to building resilience exist at the local level, The Global Adaptation Instituteª intends to develop an additional input, the ÒAdaptometerªÓ, to complement the metrics. This component would be subjective. The goal is to provide information on the awareness of national and local governments and the private sector to the sensitivity to climate change of their jurisdictions and operations. If they are aware, how do they plan to adapt? Questions addressed could include:1)	Does the society of the country understand its climate risks? 2)	How strong is the political will to address adaptation needs?3)	Are the decision makers and communities prepared to increase adaptive capacity? 4)	Is there an open, inclusive, results oriented national and sub-national dialogue on Adaptation?5)	Do budgets reflect a commitment to adaptation?Other questions will be developed dependent on the reality, level of progress and capacity of institutions on the ground to conduct such polls.The Adaptometerª would use information from the field that describes the level of knowledge, the policies in place, the allocation of budgets, and the overall commitment of society to adapting to climate change and other global trends. Concretely, polling/questionnaires would be conducted on public officials at national and local levels, plus representatives of the private sector, civil society, unions, universities, and think tanks. This additional component could provide valuable information consistent with the framework and purpose of our model. The component could be included as an additional category in Readiness, or as an independent axis (possibly combined with Adaptive Capacity). This concept can strengthen current metrics as well as contribute to the advocacy part of the InstituteÕs goals. Hopefully it would encourage and facilitate implementation of public policies that produce long term, fiscally sound solutions for vulnerable groups in society.Scaling to the localOne part of a country may be highly vulnerable while the remaining portion experiences few impacts from climate change and other global trends. More likely, a country will face several discreet challenges in different regions. Considering a countryÕs readiness, there may be significant disparities between regional governments in their ability to channel resources effectively and utilize private investment. This is not unique to developing nations. The worldÕs richest nation, the United States, is an iconic example of how a relatively ready nation with low vulnerability  can contain such disparities. The city of New Orleans, located along low-lying wetlands along the Gulf of Mexico, is both highly vulnerable and suffers from historic corruption within the city and throughout its parent state. Version 2.0 Ecosystem services ??Results & Rankings?Glossary of TermsAdaptometerª		an indicator that captures information from the field the level of awareness, the policies in place, the allocation of budgets, and the overall commitment of society to adapting to climate change and other natural disasters. The Adaptometerª will not only capture the ÒawarenessÓ of the public and government about adaptation but what actions are being formulated and implemented. Awareness		The reality at the local level. The populationÕsunderstanding of climate risks and belief that changes will increase adaptation capacity.Biophysical exposure	the level of adverse biophysical impacts due to climate change and other global forces.Fast indicator	an indicator that can change quickly, primarily due to human actions. Most readiness and adaptive capacity indicators are fast indicators.GaInª	The Global Adaptation Indexª produced by the Global Adaptation Institute (GAI). GAI is a non-profit environmental organization guided by a vision that building resilience against climate change and other global forces is a key component to sustainable development. The mission of GaInª is to enhance the worldÕs understanding of the urgency for adaptation and the support needed through private and public investments for developing countries. GaInTM is the ranking model presented in this White Paper.Medium indicators	an indicator that changes over time due to human actions. Change may take place slowly, over decades rather than years.Readiness	ability of a countryÕs private and public sectors to absorb additional investment resources and apply them effectively to increasing the resilience of communities to the effects of climate change.Readiness Matrix	a measurement of the comparative resilience of countries, plotting a countryÕs vulnerability to climate change versus its readiness to confront climate challenges.Slow indicator	an indicator, usually physical, that changes very slowly. It may not change in a lifetime. Human action has little impact on the rate of change. Socioeconomic sensitivity	the importance of a climate-sensitive sector of a country. In other words, how exposed or sensitive a country is to impacts on water, food, coastal zones, and human health. Socioeconomic adaptive capacity	the availability of economic, social, and institutional resources to cope with and adapt to the impacts of climate change in specific sectors. Though related, this differs from readiness indicators in that it measures specific actions taken to increase resilience in specific sectors, whereas Readiness measures a countryÕs ability to easily facilitate these increases in resilience. Vulnerability	a countryÕs socioeconomic exposure to biophysical impacts minus adaptive capacity.Weight	the relative value or importance an indicator is assigned in determining a countryÕs overall readiness and vulnerability score. ?References [Someone familiar with citations must review and adopt a consistent, accepted format]Bandura, Romina. 2008. A Survey of Composite Indices Measuring Country Performance: 2008 Update. UNDP. Available at: http://www.undp.org/developmentstudies/docs/indices_2008_bandura.pdfBarr, R.; Fankhauser, S. and Hamilton, K.). 2010. Adaptation Investments: a resource allocation framework. Mitigation and Adaptation Strategies Global Change. Volume 15, Number 8, 843-858. Available at: http://www.springerlink.com/content/f07067w33863x781/Brooks, Nick; Adger, Neil; Kelly, Mick. 2005. The determinants of vulnerability and adaptive capacity at the national level and the implications for adaptation. Global Environmental Change. 15. 151-163. Available at: www.uea.ac.uk/env/people/adgerwn/Brooksetal2005GEC.pdfCapacity Strengthening in climate change vulnerability and adaptation strategy assessment: exercise 2 Vulnerability Indicators (ENDA). Available at: c3d-unitar.org/c3d/userfiles/Guide_for_trainers.pdfClark, W. C., Ed. (2002). The state of the nation's ecosystems: Measuring the lands, waters, and living resources of the United States. Cambridge, UK, Cambridge Univ. Press.Climate Adaptation Knowledge Exchange. EcoAdapt. Available at: http://www.cakex.org/virtual-library/aboutClimate Change adaptation indicators for the natural environment (Natural England Commissioned Report NECRO38. 2010). Available at: http://naturalengland.etraderstores.com/NaturalEnglandShop/NECR038Climate change vulnerability and adaptation indicators. 2008. Harley, Mike; Horrocks, Lisa; and Hodgson, Nikki. Available at: http://air-climate.eionet.europa.eu/reports/ETCACC_TP_2008_9_CCvuln_adapt_indicatorsClimate Vulnerability Monitor 2010. 2010. DARA Ð Climate Vulnerability Forum. Available at: http://daraint.org/climate-vulnerability-monitor/climate-vulnerability-monitor-2010/download-the-report/Climate Research Unit. http://www.cru.uea.ac.uk/Country Policy and Institutional Assessments (CPIA), 2009 Assessment Questionnaire. 2009. World Bank. Cutter, Susan; Emrich, Christopher; Webb, Jennifer; and Daniel Morath. 2009. Social Vulnerability to Climate Variability Hazards: A Review of the Literature. Oxfam America. Available at: http://adapt.oxfamamerica.org/resources/Literature_Review.pdf Davis, K.; Kingsbury, B; and Merry, S. 2010. Indicators as technology of global governance. IILI Working Paper. Global  Administration Law Series. Available at: http://www.iilj.org/publications/2010-2.Davis-Kingsbury-Merry.aspDebels, P.; C. Szlafsztein; P. Aldunce; C. Neri; Y. Carvajal; M. Quintero-Angel; A. Celis: A. Bezanilla; D. Martinez. 2009. IUPA: a tool for the evaluation of the general usefulness of practices for adaptation to climate change and variability. Natural Hazards 0:211-233. Doing Business Index. 2010. World Bank. Available at: http://www.doingbusiness.org/rankingsEM - DAT. The International Disaster Database. Center for Research on the Epidemiology of Disasters - CRED. http://www.emdat.be/Economic Conditions Snapshot, December 2010: McKinsey Global Survey results. 2010. McKinsey and Company. Available at: http://www.mckinseyquarterly.com/Economic_Conditions_Snapshot_December_2010_McKinsey_Global_Survey_results_2720Economic Freedom of the World (EFW) Index. 2010. Fraser Institute. Available at: http://www.freetheworld.com/release.htmlEIU Business Environment Rankings. 2010. Economist Intelligence Unit. Available at: http://www.eiu.com/public/Environmental Sustainability Index. Benchmarking National Environmental Stewardship. 2005. Socioeconomic Data and Applications Center. Earth Institute, Columbia University and Yale Center for Environmental Law and Policy. Available at: http://sedac.ciesin.columbia.edu/es/esi/downloads.htmlEnvironmental Performance Index. 2010. Center for International Earth Science Information Network, Columbia Univeristy and Yale Center for Environmental Law & Policy, Yale University. Available at: http://epi.yale.edu/Environmental Vulnerability Index. 2005. South Pacific Applied Geoscience Commission (SOPAC) and United Nations Environment Programme. Available at: http://www.vulnerabilityindex.net/EVI_Background.htmFailed States Index. 2010. Foreign Policy and Fund for Peace. Available at: http://www.fundforpeace.org/web/index.php?option=com_content&task=view&id=452&Itemid=900FŸssel, Hans-Martin (2010). How Inequitable Is the Global Distribution of Responsibility, Capability, and Vulnerability to Climate Change: A comprehensive Indicator-based Assessment. Global Environmental Change 20(4):597-611, 2010 FŸssel, H. (2009) Review and Quantitative Analysis of Indices of Climate Change Exposure, Adaptive Capacity, Sensitivity, and Impacts. Background Note to World Bank Development Report 2010. Available at: http://wdronline.worldbank.org/worldbank/a/nonwdrdetail/145Global Climate Risk Index. 2011. Germanwatch and Climate Action Europe. Available at: http://www.germanwatch.org/klima/cri.htmGlobal Competitiveness Index. 2010. World Economic Forum. Available at: http://www.weforum.org/issues/global-competitivenessHedger, M.M., Horrocks, L., Mitchell, T., Leavy, J., and  Greeley, M. (2010) ÔEvaluation of adaptation to climate change from a development perspectiveÕ, IN: Van Den Berg, R. and Feinstein, O.N. (Eds), ÔEvaluating Climate Change and DevelopmentÕ, Transaction Publishers for World Bank Series on Development.Hinkel, J. (2011) ÒIndicators of vulnerability and adaptive capacity: Towards a clarification of the scienceÐpolicy interfaceÓ, Global Environmental Change, Volume 21, Issue 1, Pages 198-208. Available at: www.loceanipsl.upmc.fr/~ESCAPE/Hinkel_2011.pdfHorrocks, Lisa (AEA Group); McKenzie et al. (IDS). 2008. Institute of Development Studies. Available at: http://www.ids.ac.uk/go/idsproject/evaluating-adaptation-to-climate-change-from-a-development-perspectiveHow we can bend the curve: Global Footprint Network Annual Report. 209. Global Footprint Network. Available at: http://www.footprintnetwork.org/en/index.php/GFN/Human Development Index. 2010. UNDP. http://hdr.undp.org/en/statistics/hdi/Human Poverty Index. 2010. UNDP. Available at: http://hdr.undp.org/en/statistics/indices/hpi/ Index of Economic Freedom. 2010. The Heritage Foundation. Available at: http://www.heritage.org/index/Kaufmann, Daniel; Kraay, Aart; Mastruzzi, Massimo. World Governance Indicators. 2009. Brookings Institution; World Bank Development Economics Research Group; and World Bank Institute. Available at: http://info.worldbank.org/governance/wgi/index.aspKlein, R. 2009. Identifying Countries that are particularly vulnerable to the adverse effects of climate change: an academic of a political challenge (CCLR 3/2009 284-291. Measuring adaptation to climate change: a proposed approach. DEFRA. 2010. Available at: http://archive.defra.gov.uk/environment/climate/.../100219-measuring-adapt.pdfMitchell, Tom; van Aalst, Maarten; and Villanueva, Paula Silva. 2010. Assessing Progress on Integrating Disaster Risk Reduction and Climate Change Adaptation in Development Processes. Institute of Development Studies. Available at: community.eldis.org/.59e0d267/Convergence.pdf   Moss, Richard; Malone, Elizabeth; and Brenkert, Antoinette. 2001. Vulnerability to Climate Change: A Quantitative Approach. Pacific Northwest Laboratory for the United States Department of Energy. Available at: http://www.globalchange.umd.edu/publications/118/Myers, Mark et al. 2007. "USGS Goals for the Coming Decade.Ó Science, 12 October: Vol. 318. no. 5848, pp. 200-201. 2007Vulnerability Indices for Planning Climate Change Adaptation. 2010. National Adaptation Program of Action (NAPA). Global Environment Facility. Available at: http://www.napa-pana.org/National Aggregates of Geospatial Data Collection. Population, Landscape and Climate Estimates (PLACE). 2007. SEDAC. http://sedac.ciesin.columbia.edu/place/Stewart, Richard; Kingsbury, Benedict; and Rudyk, Bryce. 2009. Climate Finance: Regulatory and Funding Strategies for Climate Change and Global Development. New York University Abu Dhabi Institute. New York University Press. New York and London. Available at: http://www.iilj.org/climatefinance/documents/Stewartetal-ClimateFinance.pdfSullivan, Caroline et al. 2003. The Water Poverty Index: Development and application at the community scale. Natural Resources Forum. 27. 189-199. Available at: ftp://ftp.fao.org/agl/emailconf/wfe2005/narf_054.pdfSzlafsztein, Claudio; Aldunce, Paulina; and Neri, Carolina. 2008. Available at: captura.uchile.cl/jspui/bitstream/.../Evaluacion%20practicas%20utiles.pdfWater Scarcity Index. 2008. UNEP. Available at: http://www.unep.org/dewa/vitalwater/article77.htmlweADAPT 3.0. Collaborating on Climate Adaptation. Available at: http://weadapt.org/Wheeler, David. 2011. Quantifying Vulnerability to Climate Change: Implications for Adaptation Assistance Ð Working Paper 240. Center for Global Development. Available at: http://www.cgdev.org/content/publications/detail/1424759World Bank Indicators. 2010. Available at: http://data.worldbank.org/indicatorWorld Press Freedom Rating. 2010. Reporters Without Borders. Available at: http://en.rsf.org/press-freedom-index-2010,1034.htmlWorld Resources Report 2010 Framing Paper: Decision Making in a Changing Climate. World Resources Institute. Available at: www.worldresourcesreport.org/files/wrr/framing_paper.pdfWorld Telecommunication/ICT Indicators Database. 2010. International Telecommunications Union. Available at: http://itu.int/enVšršsmarty, C.J. et al. 2010. Global threats to human water security and river biodiversity. Nature 467, 555Ð561, 19 Aug. 

\end{document}  